Multi-Robot Probabilistic Mapping and Exploration

### 2018-07-05 Dissertation/Thesis/tex/ae2D.tex
tylee edit

\paragraph{Algorithm} We present an algorithm pseudocode providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}). Much like the algorithm pseudocode for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.

>>>>

\paragraph{Algorithm} We present an algorithm pseudo-code providing the necessary steps to obtain the objective function for a single measurement ray (Algorithm \ref{alg:RayExpectedEntropyGain}). Much like the algorithm pseudo-code for the ray-by-ray inverse sensor model (Algorithm \ref{alg:RayByRayISM}), the variable $a_\text{temp}$ serves as an intermediate variable designed to avoid repeated calculations.
Since this algorithm operates as a function, fixed indices and condition variables are removed for simplification.

### 2018-07-05 Dissertation/Thesis/tex/ae3Dcomplex.tex
tylee edit

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search for collision avoidance and motion planning. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidian distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.

>>>>

Next, we show how $\mathbf{m}_{\text{reduced},k}$ is used in Dijkstra's search for collision avoidance and motion planning. Defining $0<P_\text{coll}<1$ as the acceptable probability of collision, every grid cell of $ m_\text{reduced}$ is considered a safe and unvisited node if its occupancy probability, and the occupancy probabilities of its neighbors (sharing a face, edge, or corner), are below $P_\text{coll}$. A cost map is built from the starting robot location by neighboring nodes, where the cost to travel to another node is based on Euclidean distance: a cell sharing a face is $\alpha$ away, a cell sharing an edge is $\sqrt2\alpha$ away, and a cell sharing a corner is $\sqrt3\alpha$ away. Generating a cost map for all reachable cells in $\mathbf{m}_{\text{reduced},k}$ provides a collision-free travel cost for all reachable candidate pose locations. Once an optimal pose is selected, the path is easily found from the cost map using steepest descent.

### 2018-07-05 Dissertation/Thesis/tex/frontmatter.tex
tylee edit

\newglossaryentry{matrix}{
    name={matrix},
    plural={matrices},
    description={rectangular array of quantiities}
}

@@

\newglossaryentry{F}{
    type=symbols,
    name={\ensuremath{F}},
    sort=F,
    description={External Force}
}
% Some abstract text
\abstract{
This dissertation focusses on robotic mapping and exploration of uncertain environments. Computational algorithms are developed to provide important probability and uncertainty properties. These algorithms are designed for real-time implementation for robotic autonomy.

Probabilistic occupancy grid mapping is developed using Bayesian probability. A novel approach to this problem is explored, which uses important physical properties of the environment and stochastic properties of depth sensors. We develop an exact solution to occupancy grid probability that can be achieved in real-time using sensor properties and conventional assumptions of an occupancy grid. The rapid computation allows the algorithm to consider large scans of measurements in 2D and 3D environments. The mapping algorithm is demonstrated with several numerical examples and experiments.

The next topic this dissertation addresses is autonomous exploration, where a robot selects actions to maximize its knowledge about the probabilistic map. We select Shannon's entropy as a metric that represents grid cell uncertainty. Using the earlier contributions on probabilistic occupancy grid mapping, we determine the expected value of entropy change from possible future measurements. This entropy change provides important insights for where a robot should move to maximize its mapping information gain. Dijkstra's search is integral to the algorithm to account for collision-free distances during motion planning. This algorithm is designed for 2D and 3D, where computation time is carefully considered to ensure real-time algorithm performance. Several versions of the exploration algorithm are applied to simulations and experiments.

The final topic of this dissertation relates to multi-vehicle cooperative scenarios. The mapping algorithm is multi-threaded to accept measurements from multiple sources with differing sensor properties. More importantly, the exploration algorithm structure is modified with a bidding-based framework. A series of auctions determines where robots should travel such that the members act together as a team. This solution is further extended to multi-vehicle patrol, where robots begin with an uncertain map, autonomously explore the space, and periodically revisit regions. Autonomous multi-vehicle patrol is accomplished through map degradation, where the probabilistic map becomes more uncertain over time, and the robots must revisit these spaces. These complex algorithms are demonstrated with numerical simulations.

>>>>

\newglossaryentry{matrix}{
    name={matrix},
    plural={matrices},
    description={rectangular array of quantities}
}

@@

\newglossaryentry{F}{
    type=symbols,
    name={\ensuremath{F}},
    sort=F,
    description={External Force}
}
% Some abstract text
\abstract{
This dissertation focuses on robotic mapping and exploration of uncertain environments. Computational algorithms are developed to provide complete stochastic information of the environments. These algorithms are designed for real-time implementation for robotic autonomy.

First, probabilistic occupancy grid mapping is developed according to Bayesian framework. A novel approach to this problem is explored, which uses important physical properties of the environment and stochastic properties of depth sensors. We develop an exact solution to occupancy grid probability that can be achieved in real-time using sensor properties and conventional assumptions of an occupancy grid. The rapid computation allows the algorithm to consider large scans of measurements in 2D and 3D environments. The mapping algorithm is demonstrated with several numerical examples and experiments.

The next topic is autonomous exploration, where a robot selects actions to maximize its knowledge about the probabilistic map. We select Shannon's entropy as a metric that represents grid cell uncertainty. Using the earlier contributions on probabilistic occupancy grid mapping, we determine the expected value of entropy change from possible future measurements. This entropy change provides important insights for where a robot should move to maximize its mapping information gain. Dijkstra's search is integral to the algorithm to account for collision-free distances during motion planning. This algorithm is designed for 2D and 3D, where computation time is carefully considered to ensure real-time algorithm performance. Several versions of the exploration algorithm are applied to simulations and experiments.

The final topic of this dissertation relates to multi-vehicle cooperative scenarios. The mapping algorithm is revised to accept measurements from multiple sources with differing sensor properties. More importantly, the exploration algorithm structure is modified with a bidding-based framework. A series of auctions determines where robots should travel such that the members act together as a team. This solution is further extended to multi-vehicle patrol, where robots begin with an uncertain map, autonomously explore the space, and periodically revisit regions. Autonomous multi-vehicle patrol is accomplished through map degradation, where the probabilistic map becomes more uncertain over time, and the robots must revisit these spaces. These complex algorithms are demonstrated with numerical simulations.

### 2018-07-05 Dissertation/Thesis/tex/intro.tex
tylee edit

The second topic of this dissertation focusses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}. Unlike conventional motion planning, where the map of the environment is assumed known, the robot must select motions with only the current knowledge of the environment. Therefore, real-time implementation of an active motion planning scheme is necessary while the probabilistic map is changing.

@@

This dissertation is organized as follows. Chapter \ref{chap:pogm} introduces a novel approach to solve for the exact Bayesian probability of grid cells in real-time. Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a novel approach to predict future map uncertainty is presented. Next, Chapter \ref{chap:ae2D} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections. Chapter \ref{chap:multivehicle} applies this 3D mapping and exploration to multi-vehicle exploration and patrol scenarios using a bidding-based framework. Then, Chapter \ref{chap:ae3Dcomplex} extends exploration to complex 3D environments without 2D map projections. In these chapters, several numerical examples demonstrate the efficacies of the approaches. Furthermore, Chapter \ref{chap:Experiments} presents three experiments, which show the key contributions in this dissertation. Finally, these are followed by conclusions in Chapter \ref{chap:Conclusions}.

@@

% TODO: update this list

>>>>

The second topic of this dissertation focuses on the motions robots must choose to maximize their knowledge about the map while avoiding collisions, known as \emph{autonomous exploration}. Unlike conventional motion planning, where the map of the environment is assumed known, the robot must select motions with only the current knowledge of the environment. Therefore, real-time implementation of an active motion planning scheme is necessary while the probabilistic map is changing.

@@

This dissertation is organized as follows. Chapter \ref{chap:pogm} introduces a novel approach to solve for the exact Bayesian probability of grid cells in real-time. Then, Chapter \ref{chap:ae2D} extends these contributions to autonomous exploration in 2D space, where a computational approach to predict future map uncertainty is presented. Next, Chapter \ref{chap:ae2D} applies the proposed occupancy grid mapping approach to 3D environments, and presents a method to simplify autonomous exploration in 3D environments with map projections. Chapter \ref{chap:multivehicle} applies this 3D mapping and exploration to multi-vehicle exploration and patrol scenarios using a bidding-based framework. Then, Chapter \ref{chap:ae3Dcomplex} extends exploration to complex 3D environments without 2D map projections. In these chapters, several numerical examples demonstrate the efficacies of the approaches. Furthermore, Chapter \ref{chap:Experiments} presents three experiments, which show the key contributions in this dissertation. Finally, these are followed by conclusions in Chapter \ref{chap:Conclusions}.

@@

% TODO: update this list
% TL: the paper in preparation can be added as well

### 2018-07-05 Dissertation/Thesis/tex/multivehicle.tex
tylee edit

\subsection{Receding Horizon Framework}

This algorithm is further aided by following a receding horizon framework for improved information gain maximizations and collision-avoidance with dynamic obstacles and other robots. A receding horizon simply repeats the autonomous exploration steps as quickly as possible over a finite time period, frequently before a robot reaches its desired pose. Since exploration optimizations can only occur during exploration updates, a receding horizon maximizes the rate at which these updates occur. These updates require time for computation, over which the robot can move $d_\text{opt}$, used to optimize the robot travel distances with \refeqn{BumpFun}. Hence, the receding horizon framework serves to repeat optimizations as quickly as possible with maximal robotic movement while exploring a changing occupancy grid.

>>>>

\subsection{Receding Horizon Framework}
% TL: can we expand this section a little bit. for example, discussion on dynamic obstacles, or pseudo-code can be added.
This algorithm is further aided by following a receding horizon framework for improved information gain maximization and collision-avoidance with dynamic obstacles and other robots. A receding horizon simply repeats the autonomous exploration steps as quickly as possible over a finite time period, frequently before a robot reaches its desired pose. Since exploration optimizations can only occur during exploration updates, a receding horizon maximizes the rate at which these updates occur. These updates require time for computation, over which the robot can move $d_\text{opt}$, used to optimize the robot travel distances with \refeqn{BumpFun}. Hence, the receding horizon framework serves to repeat optimizations as quickly as possible with maximal robotic movement while exploring a changing occupancy grid.

### 2018-07-05 Dissertation/Thesis/tex/pogm.tex
figure captions in chap 3 updated
tylee edit

In this chapter, we present the exact solution to occupancy grid map probability. This Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities. This solution is simulated in several examples to show the efficacy of the approach.

@@

Using this, \refeqn{ThirdBayesRule} is rewritten as
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})
&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{l\in\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Fusion}
\end{align}
where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$, and $\eta_i$. Here, we introduce a new term $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for computational efficiency as
\begin{align}
&\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
\nonumber\\&=
\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber\\&\quad
+
\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t),
\end{align}
where we have used \refeqn{Unnormalized}. Similarly, its complement is
\begin{align}
P(\bar{\mathbf{m}}_i|{{X_{1:t}}},Z_{1:t})
&=\xi_i P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{l\in\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Bar_Fusion}
\end{align}
where $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
\begin{align}
&\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq\frac{\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
\nonumber\\
&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber
\\
&\quad
+\frac{
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
\end{align}
which is obtained from \refeqn{tildePbar}. Since % TODO: ref in appendix
\begin{align*}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
\end{align*}
the normalizer $\xi_i$ is obtained using \refeqn{ISM_Fusion} and \refeqn{ISM_Bar_Fusion} as,
\begin{align}
&\xi_i=
\bigg[
P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},{X_{1:t}},Z_{1:t-1})
\nonumber\\&\quad
+
P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\bigg]^{-1},\label{eqn:xi}
\end{align}
which is substituted into \refeqn{ISM_Fusion} to obtain the complete scan inverse sensor model $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.

The synergistic scan inverse sensor model algorithm is summarized with the pseudocode of Algorithm \ref{alg:SynergisticScanISM} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity. This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, defined as
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=\frac{
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
\end{align*}
as well as $d$ and $e$ for the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudocode for simplicity.

@@

In short, we propose two alternatives for combining inverse sensor models for multiple rays within a scan. The ray-by-ray approach updates the entire map one ray at a time with some dependancies among the rays, while the synergistic approach considers all rays together where rays are assumed mutually independent.

>>>>

In this chapter, we present the exact solution to occupancy probability for a given depth measurement. This Bayesian approach uses the stochastic properties of sensors, and exploits important patterns in conditional probabilities. This solution is simulated in several examples to show the efficacy of the approach.

@@

\sloppy Using this, \refeqn{ThirdBayesRule} is rewritten as
\begin{align}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})
&=\xi_i P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{l\in\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Fusion}
\end{align}
where the normalizer $\xi_i$ is composed of the product of normalizers $\zeta_i$, and $\eta_i$. Here, we introduce a new term $\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ for computational efficiency as
\begin{align}
&\hat P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq \frac{\tilde P(\mathbf{r}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\mathbf{m}_i|X_{1:t-1},Z_{1:t-1})}
\nonumber\\&=
\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber\\&\quad
+
\bigg\{\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,k+},X_t),
\end{align}
where we have used \refeqn{Unnormalized}. Similarly, its complement is
\begin{align}
P(\bar{\mathbf{m}}_i|{{X_{1:t}}},Z_{1:t})
&=\xi_i P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{l\in\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
,
\label{eqn:ISM_Bar_Fusion}
\end{align}
where $\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})$ is defined as
\begin{align}
&\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\triangleq\frac{\tilde P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})}{P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t-1})}
\nonumber\\
&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\} p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})
\nonumber
\\
&\quad
+\frac{
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
\end{align}
which is obtained from \refeqn{tildePbar}. Since % TODO: ref in appendix
\begin{align*}
P(\mathbf{m}_i|X_{1:t},Z_{1:t})+P(\bar{\mathbf{m}}_i|X_{1:t},Z_{1:t})=1,
\end{align*}
the normalizer $\xi_i$ is obtained using \refeqn{ISM_Fusion} and \refeqn{ISM_Bar_Fusion} as,
\begin{align}
&\xi_i=
\bigg[
P(\mathbf{m}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\mathbf{r}_{l,k}|z_{t,l},{X_{1:t}},Z_{1:t-1})
\nonumber\\&\quad
+
P(\bar{\mathbf{m}}_i|{X_{1:t-1}},Z_{1:t-1})
\prod_{\mathcal L_i}
\hat P(\bar{\mathbf{r}}_{l,k}|z_{t,l},X_{1:t},Z_{1:t-1})
\bigg]^{-1},\label{eqn:xi}
\end{align}
which is substituted into \refeqn{ISM_Fusion} to obtain the complete scan inverse sensor model $P(\mathbf{m}_i|X_{1:t},Z_{1:t})$.

The synergistic scan inverse sensor model algorithm is summarized with the pseudocode of Algorithm \ref{alg:SynergisticScanISM} as a computationally efficient, recursive algorithm that avoids repeated calculations of the same quantity. This algorithm utilizes the following temporary variables to develop the algorithm in an efficient recursive form, defined as
\begin{align*}
a_k&=\sum_{i=1}^{k-1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1}),
\\
b_k&=\prod_{j=0}^{k-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1}),
\nonumber\\
c_k&=\frac{
\sum_{i=k+1}^{n_{r,l}+1}\bigg\{\prod_{j=0}^{i-1}P(\bar{\mathbf{r}}_{l,j}|X_{1:t-1},Z_{1:t-1})\bigg\}p(z_{t,l}|\mathbf{r}_{l,i+},X_t)P(\mathbf{r}_{l,i}|X_{1:t-1},Z_{1:t-1})}{P(\bar{\mathbf{r}}_{l,k}|X_{1:t-1},Z_{1:t-1})},
\end{align*}
as well as $d$ and $e$ for the two terms composing \refeqn{xi}. Once again, the history of poses and measurement scans are removed from the pseudo-code for simplicity.

@@

In short, we propose two alternatives for combining inverse sensor models for multiple rays within a scan. The ray-by-ray approach updates the entire map one ray at a time with some dependencies among the rays, while the synergistic approach considers all rays together where rays are assumed mutually independent.
